
\input starlog

\centerline{\bf Starlog II}

\smallskip

\centerline{John Douglas Harold Pritchard}

\break



It had been difficult for Shevek to take the space walk.  Takver felt
his discomfort with it.  As a physicist, Shevek knew the dangers well
enough for a healthy case of hypochondria.  Takver knew the
probabilities well enough to know more about his hyper emotional body
awareness than he did.  It grew out of the earth they lived on,
another wave relation in their space.  Another subject, another
language.

\bigskip

Cornelius surveyed the orange plains of the martian riverbed through
vizov.  The life there was growing by his efforts, and those of the
PDC in raising atmospheric elements.  The terraformation of Mars would
achieve its next milestone this quarter.

\bigskip

Shevek looked down at the planetoid under him and entered the airlock.
He sat three hours.  She was here.  He exited with a copper and gold
visor, and pulled himself to the stations nadir.  Takver grabbed him
when he came over the aerodynamic bow of the structure.  She was home.

\bigskip

Cornelius would build a lake in this place, wooded with pine and oak.
It would take a hundred years, and would be named Lake Stuart.  The
next decade would develop an increasingly large scale dirt farm,
emerging from the ecosystem of the lake bed.

Last month it rained ten inches.  The clouds cooled the atmosphere at
his altitude of three hundred meters above the planned tidal mean.
When they cleared it was fifteen degrees.  A typical thunderstorm
lowered the ambient temperature by five degrees.

Stuart's farm covered the two miles of flood plane at its greatest
extent.  The hills rose to fifty meters.  Covered in grass, the plane
sprouted wild flowers, worms and wasps.  Cornelius readied for trees
and bats and birds.  This quarter he would introduce dragonflies,
ahead of the birds.

With the pines came warblers and then woodpeckers.  The PDC's seeded
species had come to bloom across Stuart's and many similar areas
across the planet.  Arid covering projects had achieved the area
necessary to the lifecycle.  From space the planet had become green
and tan and black.  Water production balanced the ground projects.
Oceans were scheduled for coming decades.

Cornelius sat at the table Anna set for him.  She prepared soup on a
wood fired stove that was molded into the hillside, into their home.
Over the decades they had constructed and reconstructed the structure
to better fit into the evolving landscape.  

Fifty years ago they had started the lake project.  The PDC's projections
were combined with a dam to add a lake to the river, and from that
foundation the heart of a wood that could survive them.

\bigskip

Cornelius was two thousand, seven hundred and thirty seven years old.
The grandson of Shevek and Takver, who lived on earth and its moon.
His lake was the center of his hemisphere.  Anna was the same age.
Friends knew Cornelius and Anna of mars and Shevek and Takver of luna
as a theme.  While Shevek and Takver were much older, and they only
rarely met, analog communications links permitted the human society on
earth, moon and mars a great deal of intercourse.

\bigskip

Shevek's reply to his message about the bird virus included a
modification to his STC.  Cornelius lifted the virus from the dead
bird with success.  He could trace environmental effects for their
sources, and with this knowledge judge the benefits of introducing
changes.  The processes were well understood.  The virii were
regulated by the PDC, but their evolutions interacted with the
environment.  They shouldn't be killing birds.

The STC analysis revealed a process operating in his wood and
elsewhere.  He identified a viral inhibitor and its possibility space.
Shevek's modifications to the STC permitted Cornelius a greater level
of confidence in the probable outcomes from releasing the genetically
modified descendents of his birds into the wild.  Shevek had said
something about orders of magnitude greater compute power, but how he
did it and what he did with it rambled into the ocean.  The birds
concerned him.  He knew that a good outcome was a no-brainer, and that
a probably good outcoome worth doing.  In this case, the outcome was
good.  

The birds would stop dying.  Under these conditions the environment
would stabilize.

\bigskip



\bye
